Do Alternate Wetting and Drying Irrigation Technology and Nitrogen Rates Affect Rice Sheath Blight?

### 2018-04-25 analysis/paper/paper.Rmd
Fix typos, add table

library(bookdown)
library(knitr)
library(kableExtra)
library(rice.awd.pests)
library(ggpubr)
```

@@

Sheath blight (ShB) disease (*Rhizoctonia solani* Kühn), anastamosis group 1 [(teleomorph: *Thanatephorus cucumeris* (A.B. Frank) Donk.] of rice is an economically important disease worldwide throughout tropical and temperate production areas. The disease can cause up to 50\% yield losses [@Marchetti1991] under conducive conditions. Studies by Savary et al. [-@Savary2000] estimated yield losses in tropical irrigated lowland rice in Asia at 5-10\%. As resistance is difficult to breed for [@Srinivasachary2011], the disease is best managed through an integrated disease management approach which includes resistance [@Kumar2009] fungicide applications where affordable [@Rush1983; @Groth2008], nitrogen management [@Slaton2003; @Castilla1996; @Tang2007] and water management [@Castilla1996].

@@

The plot design was a split-plot design with four replicates where irrigation was the main plot and nitrogen (N) rate was the split-plot treatment. The main plot treatments were alternate wetting and drying (AWD) and continuously flooded (FLD) or farmers' practice, the control treatment.

@@

The split-plot treatments were different rates of nitrogen, N~0~ (no nitrogen supply), N~1~ (100 kg per ha applied as urea in three splits at final harrowing, active tillering and panicle initiation) and N~2~ (120 kg per ha applied as urea in three splits at final harrowing, active tillering and panicle initiation).

@@

The plots were inoculated 20 days after transplanting using 151g of inoculum per plot (4m x 11m).

@@

Split-plot N rates differed from the 2015 rates with only two nitrogen treatments, N~3~ (60 kg per ha as urea split into two applications) and N~4~ (180 kg per ha in three splits) was applied.

@@

Nurseries of NSIC Rc222 were established on 7 January 2015 for the 2015 and 2016 experiments, respectively. Seedlings were transplanted on 9 January 2015 and 20 to 22 January 2016 in hills with six to eight seedlings per hill with a distance of 20 cm within and between rows.

@@

Because the data were collected on an ordinal scale, but not evenly spaced, the data are converted to the midpoint value of the percent range for severity. Area under the disease progress stairs (AUDPS) [@simko2012area] was calculated for both disease incidence and severity. As most of the data's residuals did not meet assumptions for normality, the analysis was carried out using multivariate generalised linear mixed models implemented in the `MCMCglmm` package [@MCMCglmm2010] in the R programming environment [@R2018].

@@

In 2016 the N~4~ treatment was significantly different from the N~3~ treatment. The AWD water management was also significantly different from the control flooding treatment.

@@

In 2016 the none of the treatments, nitrogen rate or water management, were significantly different from the control treatment for leaf ShB severity.

@@

@@

Table of N application timings and dates here...

>>>>

library(bookdown)
library(knitr)
library(flextable)
library(rice.awd.pests)
library(ggpubr)
```

@@

Sheath blight (ShB) disease (*Rhizoctonia solani* Kühn), anastamosis group 1 [(teleomorph: *Thanatephorus cucumeris* (A.B. Frank) Donk.] of rice is an economically important disease worldwide throughout tropical and temperate production areas. Under conducive conditions he disease can cause up to 50\% yield losses [@Marchetti1991]. Studies by Savary et al. [-@Savary2000] estimated normal yield losses in tropical irrigated lowland rice in Asia at 5-10\%. The disease is best managed through an integrated disease management approach which includes resistance [@Kumar2009, @Srinivasachary2011] fungicide applications where affordable [@Rush1983; @Groth2008], nitrogen management [@Slaton2003; @Castilla1996; @Tang2007] and water management [@Castilla1996].

@@

The plot design was a split-plot design with four replicates where irrigation was the main plot and nitrogen (N) rate was the split-plot. The main plot treatments were alternate wetting and drying (AWD) and continuously flooded (FLD) or farmers' practice as the control treatment.

@@

The split-plot treatment rates of nitrogen were N~0~ (no nitrogen supply), N~1~ (100 kg per ha applied as urea in three splits at final harrowing, active tillering and panicle initiation) and N~2~ (120 kg per ha applied as urea in three splits at final harrowing, active tillering and panicle initiation).

@@

The plots were inoculated 20 days after transplanting by spreading 151g of inoculum over the split-plot area with a 1m buffer on either end (4m x 11m).

@@

Split-plot N rates differed from the 2015 rates with only two nitrogen treatments, N~3~ (60 kg per ha as urea split into two applications) and N~4~ (180 kg per ha in three splits) being applied.

@@

Nurseries of NSIC Rc222 were established on 7 January 2016 experiments. Seedlings were transplanted by hand 20 to 22 January 2016 in hills with six to eight seedlings per hill with a distance of 20 cm within and between rows.

@@

Because the data were collected on an ordinal scale, but not evenly spaced, the data were converted to the midpoint value of the percent range for severity. Area under the disease progress stairs (AUDPS) [@simko2012area] was calculated for both disease incidence and severity. As most of the data's residuals did not meet assumptions for normality, the analysis was carried out using multivariate generalised linear mixed models implemented in the `MCMCglmm` package [@MCMCglmm2010] in the R programming environment [@R2018].

@@

In 2016 the N~4~ treatment was significantly different from the N~3~ treatment, with N~4~ severity being higher. The AWD water management, which was lower than FLD, was also significantly different.

@@

In 2016 the neither of the treatments, nitrogen rate or water management, were significantly different from the control treatments for leaf ShB severity.

@@

In both experiments we were unable to detect any significant effect of AWD on sheath blight that would cause increases in the disease, which could hinder adoption of the technology. In 2016 one measurement, tiller sheath blight severity, was significantly lower for AWD than the traditional FLD treatment, which may indicate a possible adverse effect of using AWD on leaf ShB severity.

@@

```{r N-management, }
regulartable(N_management, col_keys = c("Nitrogen treatment",
                                        "Year", 
                                        "Total N (kg/ha)",
                                        "Basal",
                                        "Tillering",
                                        "Panicle Initiation")
             )
```
